---
title: "_odds_summary_: Turning Probabilistic Estimates into Clear, Decision-Ready Insights"
output:
  prettydoc::html_pretty:
    theme: cayman #tactile, architect, leonids, hpstr
    highlight: vignette #github
    math: katex #latex
  html_document: default
  word_document: default
  pdf_document: 
    latex_engine: xelatex
description: |
  This blog is about **odds_summary** for computing odds ratio of binary and categorical models!
author:
  - name: Job Nmadu
    url: https://jobnmadu.com/
categories:
  - R
date: "2026-03-23T01:00:00Z"
tags: [RStats, odds Ratio, Percentage change, Significant variables, Data Science, Machine Learning]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy      = "styler",
                      fig.align = "center",
                      comment   = NA,
                      warning   = FALSE,
                      error     = FALSE,
                      message   = FALSE,
                      collapse  = FALSE,
                      out.width = "100%",
                      dev = "ragg_png", #<- prevent default Windows device to
                      # render plots
                      dpi       = 132,
                      echo      = TRUE)
```

# Introduction

Model tuning and estimation has evolved from simple extrapolation to sophisticated probabilistic modeling frameworks. In contemporary data science, decision-makers require more than estimates, the need for clear statements about likelihood, risk, and uncertainty are critical to correct decision-making. However, despite the advances in predictive modeling, there is a persistent limitation of making quick and robust decisions from the estimated probabilities. The estimates are often not meaningfully interpreted. This is where the odds_summary function becomes strategically important. It converts probabilistic outputs into structured summaries that directly support:

> •	decision-making 

> •	risk communication 

> •	model validation 

> •	reproducible research 

In practical terms, it turns numbers into evidence.

The function is implemented as follows: 

> odds_summary(model)  

where:

`model` An R object of estimates from models covered. For now only glm, multimon and polr models are covered.

## Implement the function

```{r}

library(Dyn4cast)
library(tidyverse)

```

### Ordered Logistic Model

```{r}
library(MASS)

options(contrasts = c("contr.treatment", "contr.poly"))
house.plr <- polr(Sat ~ Infl + Type + Cont, weights = Freq, data = housing)
modelsummary::datasummary_df(odds_summary(house.plr))
```

## glm models

```{r}
counts <- c(18,17,15,20,10,20,25,13,12)
outcome <- gl(3,1,9)
treatment <- gl(3,3)
ddc <- data.frame(treatment, outcome, counts) # showing data
glm.D93 <- glm(counts ~ ., data = ddc, family = poisson())
modelsummary::datasummary_df(odds_summary(glm.D93))

```



```{r}

anorex.1 <- glm(Postwt ~ Prewt + Treat + offset(Prewt),
                family = gaussian, data = anorexia)
modelsummary::datasummary_df(odds_summary(anorex.1))
```


```{r}
clotting <- data.frame(
  u = c(5,10,15,20,30,40,60,80,100),
  lot1 = c(118,58,42,35,27,25,21,19,18),
  lot2 = c(69,35,26,21,18,16,13,12,12))
lot1 <- glm(lot1 ~ log(u), data = clotting, family = Gamma)
modelsummary::datasummary_df(odds_summary(lot1))
```


```{r}
lot2 <- glm(lot2 ~ log(u), data = clotting, family = Gamma)
modelsummary::datasummary_df(odds_summary(lot2))
```


```{r}
x <- rnorm(100)
y <- rpois(100, exp(1+x))

lm2 <- glm(y ~ x, family = quasi(variance = "mu", link = "log"))
modelsummary::datasummary_df(odds_summary(lm2))
```


```{r}
lm3 <- glm(y ~ x, family = poisson)
modelsummary::datasummary_df(odds_summary(lm3))
```

## mlogit models

```{r}
library(mlogit)
data("Fishing", package = "mlogit")
Fish <- dfidx(Fishing, varying = 2:9, shape = "wide", choice = "mode")

## a pure "conditional" model
mml <- mlogit(mode ~ price + catch, data = Fish)
modelsummary::datasummary_df(odds_summary(mml))
```

## Multinomial Logistic model

For multinomial logistic regression, each of the measure is a data.frame, so the summary is a list of data frames. The summary is for each of the levels of the dependent variable.

```{r}
library(nnet)

tinom <- multinom(Species ~ Petal.Length + Petal.Width + Sepal.Length
                  + Sepal.Width, trace = FALSE, data = iris)

odds_summary(tinom)

```

# Conclusion

The odds_summary function represents a critical advancement in the practical interpretation of probabilistic estimates within the _Dyn4cast_
environment. Its significance lies not in computation alone, but in converting statistical output into actionable knowledge. In econometric modelling systems, prediction without interpretation is incomplete. The odds_summary function closes that gap.


# Attrition

```
cite this article as:

Nmadu J (2025). odds_summary: Turning Probabilistic Estimates into Clear, Decision-Ready Insights. https://www.jobnmadu.com/r-blog/.

To cite package 'Dyn4cast' in publications use:

  Nmadu J (2025). _Dyn4cast: Dynamic Modeling and Machine Learning
  Environment_. R package version 11.11.26,
  <https://github.com/JobNmadu/Dyn4cast>.

A BibTeX entry for LaTeX users is

  @Manual{,
    title = {_Dyn4cast: Dynamic Modeling and Machine Learning Environment_},
    author = {Job Nmadu},
    year = {2025},
    note = {R package version 11.11.26},
    url = {https://github.com/JobNmadu/Dyn4cast},
  }
```


Welcome to Data Science and Machine Learning!
